The Astropy Project

### 2017-12-01 main.tex
Edits for Sections 2.3 to 2.5, inclusive.

Central to the success of \astropy is an open environment where anybody can
contribute to the project.
However, this model leads to ``organic'' growth where different features are
implemented by different people with different programming styles and
interfaces.
Thus, \astropy has a mechanism to more formally propose significant changes to
the core package (e.g., re-writing the coordinates subpackage; \citealt{ape5}),
to plan out major new features (e.g., a new file format; \citealt{ape6}), or
institute new organization-wide policies (e.g., adopting a code of conduct;
\citealt{ape8}).
This mechanism is called ``Astropy Proposal for Enhancement'' (APE) and are
modeled after the ``Python Enhancement Proposals'' (PEP) that guide the
development of the \python programming language.
In an APE, one or more authors describe in detail the proposed changes or
additions, including a rationale for the changes, how these changes will be
implemented, and, in the case of code, what the interface will be \citep{ape1}.
The APEs are discussed and refined by the community before much work is invested
into a detailed implementation; anyone is welcome to contribute to these
discussions during the open consideration period. APEs are proposed via pull
requests on a dedicated GitHub repository\footnote{\url{https://github.com/astropy/astropy-APEs}},
and anyone can therefore read the proposed APEs and leave comments in-line.
In previous APEs a community consensus emerged and APEs are accepted and become
the basis for future work at this point.
In cases where consensus cannot be reached, the
\astropy coordination committee can decide to close the discussion and
make an executive decision based on the community input on the APE in question.

@@

A major part of the \astropy project is the concept of
``Affiliated Packages''. An affiliated package is an astronomy-related
\python package that is not part of the \astropypkg core package, but
has requested to be included as part of the \astropy project's
community. These packages support the goals and vision of \astropy of
improving code reuse, interoperability, and embracing good coding
practices such as testing and thorough documentation.

Affiliated packages contain functionality that is more specialized,
have license incompatibilities, or have external dependencies (e.g., GUI
libraries) that make these packages more suitable to be separate from the
\astropypkg core package.
Affiliated packages may also be used to develop substantial new functionality
that will eventually be incorporated into the \astropypkg core package.
New functionality benefits from having a rapid development and release cycle that is not tied to that of the \astropypkg core.
% AG: rephrased this.
%Initially separate packages with typically rapid initial development and new experimental features benefit from having a release cycle that is not tied to the

Affiliated packages are listed on the main \astropy website and advertised to the community through \astropy mailing lists, so becoming an affiliated package is a good way for new and existing packages to gain exposure while at the same time promoting \astropy's high standard for code and documentation quality. This process of listing and promoting affiliated packages is one way in which the \astropy project tries to increase code reuse in the astronomical community.

Packages can become affiliated to \astropy by applying for this status on a public mailing list. The coordination committee (\sectionname~\ref{sect:coordcom}) reviews such requests and issues recommendations for the improvement of a package where possible.

@@

\subsection{Release cycle and Long Term Support}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
% edited by Adrian PW

The \astropypkg package has a regular release schedule consisting of new significant
releases every 6 months, with bugfix releases as needed \citep{ape2}.
The major releases contain new features or any significant changes, whereas
the bugfix releases only contain fixes to code or documentation but no new
features.
Some versions are additionally designated as ``Long-term support'' (LTS)
releases, which continue to receive bugfixes for 2 years following the release
with no changes to the API\@.
The LTS versions are ideal for pipelines and other applications where API
stability is essential.
The latest LTS release (v2.0) is also the last one that supports \python 2, and
will receive bug fixes until the end of 2019 \citep{ape10}.

The version numbering of the \astropypkg core package reflects this release
scheme: the core package version number uses the form x.y.z, where ``x'' is
advanced for LTS releases, ``y'' is advanced for non-LTS feature releases, and
``z'' is advanced for bugfix releases.

>>>>

Central to the success of \astropy is an open environment where anybody can
contribute to the project.
This model leads to an ``organic'' growth, where different features are
implemented by different people with different programming styles and
interfaces.
Thus, \astropy has a mechanism to more formally propose significant changes to
the core package (e.g., re-writing the coordinates subpackage; \citealt{ape5}),
to plan out major new features (e.g., a new file format; \citealt{ape6}), or
institute new organization-wide policies (e.g., adopting a code of conduct;
\citealt{ape8}).
This mechanism is called ``Astropy Proposal for Enhancement'' (APE) and are
modeled after the ``Python Enhancement Proposals'' (PEP) that guide the
development of the \python programming language.
In an APE, one or more authors describe in detail the proposed changes or
additions, including a rationale for the changes, how these changes will be
implemented, and in the case of code, what the interface will be \citep{ape1}.
The APEs are discussed and refined by the community before much work is invested
into a detailed implementation; anyone is welcome to contribute to these
discussions during the open consideration period. APEs are proposed via pull
requests on a dedicated GitHub repository\footnote{\url{https://github.com/astropy/astropy-APEs}};
anyone can therefore read the proposed APEs and leave in-line comments.
When a community consensus emerge, the APEs are accepted and become
the basis for future work.
In cases where consensus cannot be reached, the
\astropy coordination committee may decide to close the discussion and
make an executive decision based on the community input on the APE in question.

@@

A major part of the \astropy project is the concept of
``Affiliated Packages''. An affiliated package is an astronomy-related
\python package that is not part of the \astropypkg core package, but
has requested to be included as part of the \astropy project's
community. These packages support the goals and vision of \astropy of
improving code re-use, interoperability, and embracing good coding
practices such as testing and thorough documentation.

Affiliated packages contain functionality that is more specialized,
have license incompatibilities, or have external dependencies (e.g., GUI
libraries) that make these packages more suitable to be separate from the
\astropypkg core package.
Affiliated packages may also be used to develop substantial new functionality
that will eventually be incorporated into the \astropypkg core package
(e.g., \texttt{wcsaxes}).
New functionality benefits from having a rapid development and release cycle that is not tied to that of the \astropypkg core (\sectionname~\ref{sect:releasecycle}).
% AG: rephrased this.
%Initially separate packages with typically rapid initial development and new experimental features benefit from having a release cycle that is not tied to the

Affiliated packages are listed on the main \astropy website and advertised to the community through \astropy mailing lists. Therefore, becoming an affiliated package is a good way for new and existing packages to gain exposure while promoting \astropy's high standard for code and documentation quality. This process of listing and promoting affiliated packages is one way in which the \astropy project tries to increase code re-use in the astronomical community.

Packages can become affiliated to \astropy by applying for this status on a public mailing list. The coordination committee (\sectionname~\ref{sect:coordcom}) reviews such requests and issues recommendations for the improvement of a package, where applicable.

@@

\subsection{Release cycle and Long Term Support}
\label{sect:releasecycle}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
% edited by Adrian PW

The \astropypkg package has a regular release schedule consisting of new significant
releases every 6 months, with bugfix releases as needed \citep{ape2}.
The major releases contain new features or any significant changes, whereas
the bugfix releases only contain fixes to code or documentation but no new
features.
Some versions are additionally designated as ``Long-term support'' (LTS)
releases, which continue to receive bug fixes for 2 years following the release
with no changes to the API\@.
The LTS versions are ideal for pipelines and other applications, where API
stability is essential.
The latest LTS release (v2.0) is also the last one that supports \python 2; It
will receive bug fixes until the end of 2019 \citep{ape10}.

The version numbering of the \astropypkg core package reflects this release
scheme: the core package version number uses the form \texttt{x.y.z}, where ``x'' is
advanced for LTS releases, ``y'' for non-LTS feature releases, and
``z'' for bugfix releases.
